Sequence Specific Modeling of Cell-Free Protein Synthesis

### 2017-03-09 Paper_v1.tex
minor updates

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
In this study, we used sequence specific constraints based modeling to evaluate the performance of synthetic circuits in an \emph{E.~coli} cell-free protein synthesis system.
A core \emph{E.~coli} metabolic model, consisting of glycolysis, pentose phosphate pathway, amino acid biosynthesis and degradation and energy metabolism, was then augmented with sequence specific descriptions of genetic circuits which included mechanistic models of promoter function, transcription and translation.
Thus, unlike other synthetic biology modeling efforts, sequence specific constraints based modeling explicitly couples the transcription and translation of circuit components with the availability of metabolic resources.
Model parameters for transcription and translation were taken from literature, allowing a first principles prediction of circuit performance.
We tested this approach by first simulating T7 induced chloramphenicol acetyltransferase production and $\sigma_{70}$-induced deGFP expression; we then expanded these studies for a range of different proteins.
First principles predictions of circuit performance were consistent with measurements for a variety of cases.
Further, global sensitivity analysis identified the key metabolic processes that controlled circuit performance in terms of productivity, energy efficiency, and carbon yield.
A sufficient energy supply with oxidative phosphorylation is instrumental for high energy efficiency and carbon yields; in addition, the translation rate could be optimized for higher productivity.
Taken together, sequence specific constraints based modeling offers a novel means to \emph{a~priori} estimate the performance of cell-free synthetic circuits.
\end{abstract}

@@

Stoichiometric reconstructions of microbial metabolism, popularized by flux balance analysis (FBA), have become standard tools to interrogate metabolism \cite{2012_lewis_palsson_NatRevMicrobio}.
Since the first genome scale stoichiometric model of \textit{E.~coli} \cite{2000_edwards_palsson_PNAS}, stoichiometric reconstructions of hundreds of organisms, including industrially important prokaryotes such as \textit{E.~coli} \cite{Feist:2007aa} or \textit{B. subtilis} \cite{Oh:2007aa}, are now available \cite{2009_feist_palsson_NatRevMicrobio}.
In this study, we used sequence specific constraints based modeling to evaluate the performance of \emph{E.~coli} cell-free protein synthesis (CFPS).
A core \emph{E.~coli} cell-free metabolic model was developed from literature \cite{Feist:2007aa}.
This model, which described glycolysis, pentose phosphate pathway, amino acid biosynthesis and degradation and energy metabolism, was then augmented with
sequence specific descriptions of promoter function, transcription and translation processes.
Thus, sequence specific constraints based modeling explicitly coupled transcription and translation with the availability of metabolic resources.
We tested this approach by simulating the production of two model proteins, and then investigated the productivity and carbon yield for eight different proteins.
From this, higher carbon number proteins typically had lower productivity rates and carbon yields than that of the lower carbon number proteins.
Further, global sensitivity analysis identified the key metabolic processes that controlled circuit performance, showing oxidative phosphorylation as instrumental for maintaining a high carbon yield and the translation rate for productivity.
Taken together, sequence specific constraints based modeling offers a novel means to \emph{a~priori} estimate the performance of cell-free synthetic circuits.

@@

\subsection{Model derivation and validation}
The cell free stoichiometric network was constructed by removing growth associated reactions from the \textit{i}AF1260 reconstruction of K-12 MG1655 \textit{E.~coli} \cite{Feist:2007aa}.
We then added the transcription and translation template reactions of Allen and Palsson for the specific proteins of interest \cite{Allen:2003aa}.
A schematic of the metabolic network, which consisted of 272 reactions and 145 species, is shown in Fig.~\ref{fig:network}A.
Using this network, in combination with detailed promoter models, and literature values for cell free culture parameters (Table ZZ),
we simulated the sequence specific production of two model proteins, chloramphenicol acetyltransferase (CAT) and dual emission green fluorescent protein (deGFP) using
different cell-free \textit{E.~coli} extracts. The cell free metabolic network, model parameters, model code,
and each protein sequence are available in the supplemental materials.

Cell free simulations predicted CAT and deGFP production for the duration of the CFPS batch reactions (Fig. ~\ref{fig:network}B and C).
Chloramphenicol acetyltransferase (CAT), was produced under a T7 promoter in a glucose/NMP cell-free system \cite{2005_calhoun_BiotechnologyProgress} for 1 hour using glucose as a carbon and energy source (Fig.~\ref{fig:network}B).
With the exception of the first 10-15 min, the cell free prediction of CAT abundance was consistent with the measured values.
On the other hand, deGFP was produced under a P70 promoter in TXTL 2.0 \textit{E.~coli} extract for 8 hours using maltose as a carbon and energy source (Fig.~\ref{fig:network}C).
The cell free simulation captured the overall trend of deGFP abundance, but was not able to capture saturation at the end of the CFPS culture.
Uncertainty in experimental factors such as the concentration of RNA polymerase, and ribosomes,
transcription and translation elongation rates and the upper bounds for oxygen and glucose consumption rates did not alter the qualitative performance of the model.
Thus, the metabolic network and molecular description of transcription and translation were consistent with experimental measurements.

Next, we predicted deGFP production as a function of plasmid concentration (Fig.~\ref{fig:network}D).
Concentration of deGFP at each plasmid concentration was calculated by multiplying the flux of deGFP synthesis by the active time of production, approximately 8 hours in TXTL 2.0 \cite{Garamella:2016aa}.
The mean of the ensemble shows a good prediction against the measured deGFP levels, even though it under predicted deGFP concentration at the saturating point of 5 nM of plasmid concentration.
However, the ensemble and the mean of the ensemble captured the overall saturating dynamics of deGFP production as a function of plasmid concentration.
\begin{figure}[t!]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/ssFBA_network.pdf}
\caption{Sequence specific flux balance analysis. A. Core metabolic network with glycolysis, pentose phoshate pathway, TCA cycle, Entner-Doudoroff pathway. Thick gray arrows indicate withdrawal of precursors for amino acid synthesis. B. CAT production under a T7 promoter in CFPS \textit{E.~coli} extract for 1 h under glucose consumption. Error bars denote the standard deviation of experimental measurements. C. deGFP production under a P70 promoter in TXTL 2.0 \textit{E.~coli} extract for 8 h under maltose consumption. Error bars denote a 10\% coefficient of variation. D. Predicted deGFP concentration at different plasmid concentrations versus measurements of deGFP synthesized in TXTL 2.0. 95\% CI (blue region) over the ensemble of 100 sets, mean of the ensemble (black line), and experimental measurements (dots).}
\label{fig:network}
\end{figure}
These results validated our mathematical framework to model CFPS systems and predict the production of two proteins with no adjustable parameters.
It also showed that the sequence specific reactions were sufficient to predict the production of two different proteins under different promoters and cell-free systems.
Since the model accurately predicted protein production, we used our mathematical framework to understand the performance limits of CFPS.

\subsection{Analysis of CFPS performance}
Our next goal was to examine the performance of CFPS for eight different proteins under three different cases.
Each of the proteins was produced under a P70 promoter, except for CAT which was produced under a T7 promoter.
In all cases, CFPS was supplied with glucose.
In the first case, CFPS was supplied with amino acids, and the system was allowed to synthesize amino acids (AA uptake and synthesis).
In the second case, CFPS was supplied with amino acids, but the amino acid synthesis reactions were turned off (AA uptake w/o synthesis).
These amino acid synthesis reactions were blocked since during the cell-free extract preparation the cells are often supplied with amino acids; thus, the enzymes responsible for amino acid synthesis would not be present.
In the third case, CFPS was not supplied with amino acids, but the system could synthesize them (AA synthesis w/o uptake).
Eight different proteins, ranging in size, were selected to evaluate CFPS performance: bone morphogenetic protein 10 (BMP10), chloramphenicol acetyltransferase (CAT), caspase 9 (CASP9), dual emission green fluorescent protein (deGFP), prothrombin (FII), coagulation factor X (FX), fibroblast growth factor 21 (FGF21), and single chain variable fragment R4 (scFvR4).
We used ssFBA to estimate the productivity for each of these proteins for each case (Fig.~\ref{fig:Prod}).
An additional case was considered for CAT, since a comprehensive dataset is available \cite{2005_calhoun_BiotechnologyProgress}; in this case, fluxes were constrained to experimental measurements where avaliable, with the exception of CAT production which was determined by the transcription/translation parameters.

\subsubsection{Productivity}
We evaluated CFPS productivity for all eight proteins and for each case.
All cases had very similar performance for each protein and were within a standard deviation of each other (Fig.~\ref{fig:Prod}A).
%with the second case (without amino acid synthesis) having a slightly higher productivity for most proteins.
The model framework is setup to optimize for the production of each protein and is constrained by the translation rate.
This shows the system had sufficient substrates and metabolic precursors to power CFPS and synthesize each protein of interest with the same productivity, regardless of the case.
However, each protein had a different level of productivity.
For instance, BMP10 had a productivity of about 2.5 $\mu$M/h whereas CAT had a productivity of about 12 $\mu$M/h.
To examine this further, the mean productivity was plotted against the carbon number of each protein (Fig.~\ref{fig:Prod}B).
The proteins with the highest productivity had the lowest carbon number, whereas proteins with low productivity had higher carbon numbers.
This inverse trend was due to the fact that larger proteins require more amino acids and substrates to assemble them, resulting in lower productivity given the same resources.
A single trendline for all cases shows the expected productivity in CFPS depending on the carbon number of the protein of interest (available in the Supporting Information).
CAT was an outlier for the trendline, even though it was in the same order of magnitude as the trendline.
The difference in the T7 and  $\sigma_{70}$ promoters did not alter the qualitative productivity performance of CAT.
The higher productivity of CAT compared to all other proteins was most likely due to the lower transcription requirement of cytidine triphosphate which allowed a higher flux for translation.
% a further study on the nucleotide and amino acid requirements of each protein and its effect on CFPS performance should be investigated.
\begin{figure}[t!]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/Productivity.pdf}
\caption{CFPS productivity of eight proteins for four cases. Amino acid uptake and synthesis (black), AA uptake without synthesis (dark grey), AA synthesis without uptake (light grey), and constrained by experimental measurements, for CAT only (white). A. Productivity across the ensemble (error bars represent 95\% CI). B. Mean productivity versus carbon number. Single trendline (dotted line) calculated across all cases (R\textsuperscript{2} = 0.99). Asterisk: protein excluded from trendline; dagger: constrained by experimental measurements and excluded from trendline.}
\label{fig:Prod}
\end{figure}
The drop in productivity in the third case is due to the fact that glucose is the only substrate supplied and must be used to provide the necessary energy requirements for transcription and translation, as well as synthesize each amino acid required for the production of the protein of interest.

@@

The standard protein synthesis reaction was conducted according to the PANOxSP protocol with slight modifications from that described previously \cite{BIT:BIT20026}.
Unless otherwise noted, all reagents were purchased from Sigma (St. Louis, MO).
The initial mixture includes 1.2 mM ATP; 0.85 mM each of GTP, UTP, and CTP; 30 mM phosphoenolpyruvate (Roche, Indianapolis IN); 130 mM potassium glutamate; 10 mM ammonium glutamate; 16 mM magnesium glutamate; 50 mM HEPES-KOH buffer (pH 7.5); 1.5 mM spermidine; 1.0 mM putrescine; 34 $\mu$g/mL folinic acid; 170.6 $\mu$g/mL E. coli tRNA mixture (Roche, Indianapolis IN); 13.3 $\mu$g/mL pK7CAT plasmid; 100 $\mu$g/mL T7 RNA polymerase; 20 unlabeled amino acids at 2 mM each; 5 $\mu$M l-[U-\textsuperscript{14}C]-leucine (Amersham Pharmacia, Uppsala Sweden); 0.33 mM nicotinamide adenine dinucleotide (NAD); 0.26 mM coenzyme A (CoA); 2.7 mM sodium oxalate; and 0.24 volumes of E. coli S30 extract.
This reaction was modified for the energy source used such that glucose reactions have 30 mM glucose in place of PEP, and glutamate reactions do not have a secondary energy source added since glutamate is already present in high concentrations from the reaction salts. Sodium oxalate was not added since it has a detrimental effect on protein synthesis and ATP concentrations when using glucose or an early glycolytic intermediate as the energy source \cite{BIT:BIT1121}. The HEPES buffer (pKa $\sim$ 7.5) was replaced with Bis-Tris (pKa $\sim$ 6.5). In addition, the magnesium glutamate concentration was reduced to 8 mM for the glucose and glutamate reactions since a lower magnesium optimum is found when using a nonphosphorylated energy source \cite{BIT:BIT20026}. Finally, 10 mM phosphate was added to some reactions as described below in the form of potassium phosphate dibasic adjusted to pH 7.2 with acetic acid.

@@

The bounds on the transcription rate ($\mathcal{L}_{TX}=w_{TX}=\mathcal{U}_{TX}$) were modeled as:
\begin{equation}
	w_{TX} = V_{TX}^{max}\left(\frac{G}{K_{TX}+G}\right)
\end{equation}
where $G$ denotes the gene concentration and $K_{TX}$ denotes a transcription saturation coefficient.
The maximum rate of transcription $V_{TX}^{max}$ was formulated as:
\begin{equation}
	V_{TX}^{max} \equiv \left[R_{T}\left(\frac{v_{TX}}{l_{G}}\right)u\left(\kappa\right)\right]
\end{equation}
The term $R_{T}$ denotes the RNA polymerase concentration (nM),
$v_{TX}$ denotes the RNA polymerase elongation rate (nt/h),
$l_{G}$ denotes the gene length in nucleotides (nt).
The term $u\left(\kappa\right)$ (dimensionless, $0\leq u\left(\kappa\right)\leq 1$) describes a model of promoter activity, where $\kappa$ denotes promoter specific parameters.
The forms for the promoter models were taken from Moon $\emph{et~al.}$ \cite{Moon:2012ab}.
In this study, we considered two promoters: T7 and $\sigma_{70}$.
The promoter function for the T7 promoter, $u_{T7}$, was given by:
\begin{equation}
	u_{T7} = \frac{K_{T7}}{1 + K_{T7}}
\end{equation}
where $K_{T7}$ denotes a T7 RNA polymerase binding constant.
The $\sigma_{70}$ promoter function $u_{\sigma_{70}}$ (which was used for all other proteins) was formulated as:
\begin{equation}
	u_{\sigma_{70}} = \frac{K_{1}+K_{2}f_{\sigma_{70}}}{1 + K_{1}+K_{2}f_{\sigma_{70}}}
\end{equation}
where $K_{1}$ denotes the weight of RNA polymerase binding alone,
$K_{2}$ denotes the weight of RNAP-$\sigma_{70}$ bound to the promoter,
and $f_{p70}$ denotes the fraction of the $\sigma_{70}$ transcription factor bound to RNAP (modeled as a Hill function).
The values for $K_{T7}, K_{1}$, $K_{2}$ and $f_{\sigma_{70}}$, along with all other promoter parameters, are given in Table ZZ.

The translation rate ($w_{TL}$) was bounded by:
 \begin{equation}
	0\leq w_{TL} \leq V_{TL}^{max}\left(\frac{mRNA_{SS}}{K_{TL}+mRNA_{SS}}\right)
\end{equation}
where $\rm mRNA_{SS}$ denotes the steady state mRNA abundance and $K_{TL}$ denotes the translation saturation constant.
The maximum translation rate $V_{TL}^{max}$ was formulated as:
\begin{equation}
	V_{TL}^{max} \equiv \left[K_{P} R_{X}\left(\frac{v_{TL}}{l_{P}}\right)\right]
\end{equation}
The term $K_{P}$ denotes the polysome amplification constant,
$v_{TL}$ denotes the ribosome elongation rate (amino acids per hour),
$l_{P}$ denotes the number of amino acids in the protein of interest,
and $\rm mRNA_{SS}$ denotes the steady-state mRNA concentration:
\begin{equation}
	 mRNA_{SS}\simeq\frac{w_{TX}}{\lambda}
\end{equation}
where $\lambda$ denotes the rate constant controlling the mRNA degradation rate.
All translation parameters are given in Table ZZ.

\subsection*{Calculation of energy efficiency.}
Energy efficiency ($\mathcal{E}$) was calculated as the ratio of protein production to glucose consumption, both in terms of equivalent ATP molecules:
\begin{equation}\label{eqn:energy-efficiency-definition}
	\mathcal{E}=\frac{\Delta\texttt{POI}\cdot \left(2\left(\rm {ATP}_{TX}+\rm {CTP}_{TX}+\rm {GTP}_{TX}+\rm {UTP}_{TX}\right)+\rm2\cdot\rm {ATP}_{TL}+\rm {GTP}_{TL}\right)}{\Delta\texttt{GLC}\cdot\texttt{ATP}_{GLC}}
\end{equation}
where $\Delta\texttt{POI}=w_{TX}$ denotes the translation rate for the protein of interest, $\rm {ATP}_{TX}$, $\rm {CTP}_{TX}$, $\rm {GTP}_{TX}$, $\rm {UTP}_{TX}$ denote the stoichiometric coefficients of each energy species for the transcription of the protein of interest, $\rm {ATP}_{TL}$, $\rm {GTP}_{TL}$ denote the stoichiometric coefficients of ATP and GTP for the translation of the protein of interest, $\Delta\texttt{GLC}=w_{GLC}$ denotes the glucose flux, and $\rm {ATP}_{GLC}$ denotes the equivalent ATP number for glucose.
The energy species stoichiometric coefficients are available in the Supporting Information.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
In this study, we used sequence specific constraints based modeling to evaluate the performance of synthetic circuits in an \emph{E.~coli} cell-free protein synthesis system.
A core \emph{E.~coli} metabolic model, consisting of glycolysis, pentose phosphate pathway, amino acid biosynthesis and degradation and energy metabolism, was then augmented with sequence specific descriptions of genetic circuits which included mechanistic models of promoter function, transcription and translation.
Thus, unlike other synthetic biology modeling efforts, sequence specific constraints based modeling explicitly couples the transcription and translation of circuit components with the availability of metabolic resources.
Model parameters for transcription and translation were taken from literature, allowing a first principles prediction of circuit performance.
We tested this approach by first simulating T7 induced chloramphenicol acetyltransferase production and P70a-induced deGFP expression; we then expanded these studies for a range of different proteins.
First principles predictions of circuit performance were consistent with measurements for a variety of cases.
Further, global sensitivity analysis identified the key metabolic processes that controlled circuit performance in terms of productivity, energy efficiency, and carbon yield.
A sufficient energy supply with oxidative phosphorylation is instrumental for high energy efficiency and carbon yields; in addition, the translation rate could be optimized for higher productivity.
Taken together, sequence specific constraints based modeling offers a novel means to \emph{a~priori} estimate the performance of cell-free synthetic circuits.
\end{abstract}

@@

Stoichiometric reconstructions of microbial metabolism, popularized by flux balance analysis (FBA), have become standard tools to interrogate metabolism \cite{2012_lewis_palsson_NatRevMicrobio}.
Since the first genome scale stoichiometric model of \textit{E.~coli} \cite{2000_edwards_palsson_PNAS}, stoichiometric reconstructions of hundreds of organisms, including industrially important prokaryotes such as \textit{E.~coli} \cite{Feist:2007aa} or \textit{B. subtilis} \cite{Oh:2007aa}, are now available \cite{2009_feist_palsson_NatRevMicrobio}.
In this study, we used sequence specific constraints based modeling to evaluate the performance of \emph{E.~coli} cell-free protein synthesis (CFPS).
A core \emph{E.~coli} cell-free metabolic model was developed from literature \cite{Feist:2007aa}.
This model, which described glycolysis, pentose phosphate pathway, amino acid biosynthesis and degradation and energy metabolism, was then augmented with
sequence specific descriptions of promoter function, transcription and translation processes.
Thus, sequence specific constraints based modeling explicitly coupled transcription and translation with the availability of metabolic resources.
We tested this approach by simulating the production of two model proteins, and then investigated the productivity, energy efficiency, and carbon yield for eight different proteins.
From this, higher carbon number proteins typically had lower productivity rates, energy efficiency, and carbon yields than that of the lower carbon number proteins.
Further, global sensitivity analysis identified the key metabolic processes that controlled circuit performance, showing oxidative phosphorylation as instrumental for maintaining a high energy efficiency and carbon yield and the translation rate for productivity.
Taken together, sequence specific constraints based modeling offers a novel means to \emph{a~priori} estimate the performance of cell-free synthetic circuits.

@@

\subsection{Model derivation and validation}
The cell-free stoichiometric network was constructed by removing growth associated reactions from the \textit{i}AF1260 reconstruction of K-12 MG1655 \textit{E.~coli} \cite{Feist:2007aa}.
We then added the transcription and translation template reactions of Allen and Palsson for the specific proteins of interest \cite{Allen:2003aa}.
A schematic of the metabolic network, which consisted of 270 reactions and 145 species, is shown in Fig.~\ref{fig:network}A.
Using this network, in combination with detailed promoter models, and literature values for cell-free culture parameters (Table ~\ref{tbl:parameters}),
we simulated the sequence specific production of two model proteins, chloramphenicol acetyltransferase (CAT) and dual emission green fluorescent protein (deGFP) using
different cell-free \textit{E.~coli} extracts. The cell-free metabolic network, model parameters, model code,
and each protein sequence are available in the supplemental materials.

Cell-free simulations predicted CAT and deGFP production for the duration of the CFPS batch reactions (Fig. ~\ref{fig:network}B and C).
Chloramphenicol acetyltransferase (CAT), was produced under a T7 promoter in a glucose/NMP cell-free system \cite{2005_calhoun_BiotechnologyProgress} for 1 hour using glucose as a carbon and energy source (Fig.~\ref{fig:network}B).
With the exception of the first 10-15 min, the cell-free prediction of CAT abundance was consistent with the measured values.
On the other hand, deGFP was produced under a P70a promoter in TXTL 2.0 \textit{E.~coli} extract for 8 hours using maltose as a carbon and energy source (Fig.~\ref{fig:network}C).
The cell-free simulation captured the overall trend of deGFP abundance, but was not able to capture saturation at the end of the CFPS culture.
Uncertainty in experimental factors such as the concentration of RNA polymerase, ribosomes,
transcription and translation elongation rates and the upper bounds for oxygen and glucose consumption rates did not alter the qualitative performance of the model.
Thus, the metabolic network and molecular description of transcription and translation were consistent with experimental measurements.

Next, we predicted deGFP production as a function of plasmid concentration (Fig.~\ref{fig:network}D).
Concentration of deGFP at each plasmid concentration was calculated by multiplying the flux of deGFP synthesis by the active time of production, approximately 8 hours in TXTL 2.0 \cite{Garamella:2016aa}.
The mean of the ensemble shows a good prediction against the measured deGFP levels, even though it under predicted deGFP concentration at the saturating point of 5 nM of plasmid concentration.
However, the ensemble and the mean of the ensemble captured the overall saturating dynamics of deGFP production as a function of plasmid concentration.
\begin{figure}[t!]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/ssFBA_network.pdf}
\caption{Sequence specific flux balance analysis. A. Core metabolic network with glycolysis, pentose phoshate pathway, TCA cycle, Entner-Doudoroff pathway. Thick gray arrows indicate withdrawal of precursors for amino acid synthesis. B. CAT production under a T7 promoter in CFPS \textit{E.~coli} extract for 1 h under glucose consumption. Error bars denote the standard deviation of experimental measurements. C. deGFP production under a P70 promoter in TXTL 2.0 \textit{E.~coli} extract for 8 h under maltose consumption. Error bars denote a 10\% coefficient of variation. D. Predicted deGFP concentration at different plasmid concentrations versus measurements of deGFP synthesized in TXTL 2.0. 95\% CI (blue region) over the ensemble of 100 sets, mean of the ensemble (black line), and experimental measurements (dots).}
\label{fig:network}
\end{figure}
These results validated our mathematical framework to model CFPS systems and predict the production of two proteins with very few adjustable parameters.
It also showed that the sequence specific reactions were sufficient to predict the production of two different proteins under different promoters and cell-free systems.
Since the model accurately predicted protein production, we used our mathematical framework to understand the performance limits of CFPS.

\subsection{Analysis of CFPS performance}
Our next goal was to examine the performance of CFPS for eight different proteins under three different cases.
Each of the proteins was produced under a P70a promoter, except for CAT which was produced under a T7 promoter.
In all cases, CFPS was supplied with glucose.
In the first case, CFPS was supplied with amino acids, and the system was allowed to synthesize amino acids (AA uptake and synthesis).
In the second case, CFPS was supplied with amino acids, but the amino acid synthesis reactions were turned off (AA uptake w/o synthesis).
These amino acid synthesis reactions were blocked since during the cell-free extract preparation the cells are often supplied with amino acids; thus, the enzymes responsible for amino acid synthesis would not be present.
In the third case, CFPS was not supplied with amino acids, but the system could synthesize them (AA synthesis w/o uptake).
Eight different proteins, ranging in size, were selected to evaluate CFPS performance: bone morphogenetic protein 10 (BMP10), chloramphenicol acetyltransferase (CAT), caspase 9 (CASP9), dual emission green fluorescent protein (deGFP), prothrombin (FII), coagulation factor X (FX), fibroblast growth factor 21 (FGF21), and single chain variable fragment R4 (scFvR4).
We used ssFBA to estimate the productivity, energy efficiency, and carbon yield for each of these proteins for each case.
An additional case was considered for CAT, since a comprehensive dataset is available (Supporting Information); in this case, fluxes were constrained to experimental measurements where avaliable, with the exception of CAT production which was determined by the transcription/translation parameters.

\subsubsection{Productivity}
We evaluated CFPS productivity for all eight proteins and for each case.
All cases had very similar performance for each protein and were within a standard deviation of each other (Fig.~\ref{fig:Prod}A).
%with the second case (without amino acid synthesis) having a slightly higher productivity for most proteins.
The model framework is setup to optimize for the production of each protein and is constrained by the translation rate.
This shows the system had sufficient substrates and metabolic precursors to power CFPS and synthesize each protein of interest with the same productivity, regardless of the case.
However, each protein had a different level of productivity.
For instance, BMP10 had a productivity of about 2.5 $\mu$M/h whereas CAT had a productivity of about 12 $\mu$M/h.
To examine this further, the mean productivity was plotted against the carbon number of each protein (Fig.~\ref{fig:Prod}B).
The proteins with the highest productivity had the lowest carbon number, whereas proteins with low productivity had higher carbon numbers.
This inverse trend was due to the fact that larger proteins require more amino acids and substrates to assemble them, resulting in lower productivity given the same resources.
A single trendline for all cases shows the expected productivity in CFPS depending on the carbon number of the protein of interest (available in the Supporting Information).
CAT was an outlier for the trendline, even though it was in the same order of magnitude as the trendline.
The difference in the T7 and  P70a promoters did not alter the qualitative productivity performance of CAT.
The higher productivity of CAT compared to all other proteins was most likely due to the lower transcription requirement of cytidine triphosphate which allowed a higher flux for translation.
% a further study on the nucleotide and amino acid requirements of each protein and its effect on CFPS performance should be investigated.
\begin{figure}[t!]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/Productivity.pdf}
\caption{CFPS productivity of eight proteins for four cases. Amino acid uptake and synthesis (black), AA uptake without synthesis (dark grey), AA synthesis without uptake (light grey), and constrained by experimental measurements, for CAT only (white). A. Productivity across the ensemble (error bars represent 95\% CI). B. Mean productivity versus carbon number. Single trendline (dotted line) calculated across all cases (R\textsuperscript{2} = 0.99). Asterisk: protein excluded from trendline; dagger: constrained by experimental measurements and excluded from trendline.}
\label{fig:Prod}
\end{figure}
The drop in productivity in the third case is due to the fact that glucose is the only substrate supplied and must be used to provide the necessary energy requirements for transcription and translation, as well as synthesize each amino acid required for the production of the protein of interest.

@@

The standard protein synthesis reaction was conducted according to the PANOxSP protocol with slight modifications from that described previously \cite{BIT:BIT20026}.
Unless otherwise noted, all reagents were purchased from Sigma (St. Louis, MO).
The initial mixture includes 1.2 mM ATP; 0.85 mM each of GTP, UTP, and CTP; 30 mM phosphoenolpyruvate (Roche, Indianapolis IN); 130 mM potassium glutamate; 10 mM ammonium glutamate; 16 mM magnesium glutamate; 50 mM HEPES-KOH buffer (pH 7.5); 1.5 mM spermidine; 1.0 mM putrescine; 34 $\mu$g/mL folinic acid; 170.6 $\mu$g/mL E. coli tRNA mixture (Roche, Indianapolis IN); 13.3 $\mu$g/mL pK7CAT plasmid; 100 $\mu$g/mL T7 RNA polymerase; 20 unlabeled amino acids at 2-3 mM each; 5 $\mu$M l-[U-\textsuperscript{14}C]-leucine (Amersham Pharmacia, Uppsala Sweden); 0.33 mM nicotinamide adenine dinucleotide (NAD); 0.26 mM coenzyme A (CoA); 2.7 mM sodium oxalate; and 0.24 volumes of E. coli S30 extract.
This reaction was modified for the energy source used such that glucose reactions have 30-40 mM glucose in place of PEP, and glutamate reactions do not have a secondary energy source added since glutamate is already present in high concentrations from the reaction salts. Sodium oxalate was not added since it has a detrimental effect on protein synthesis and ATP concentrations when using glucose or an early glycolytic intermediate as the energy source \cite{BIT:BIT1121}. The HEPES buffer (pKa $\sim$ 7.5) was replaced with Bis-Tris (pKa $\sim$ 6.5). In addition, the magnesium glutamate concentration was reduced to 8 mM for the glucose and glutamate reactions since a lower magnesium optimum is found when using a nonphosphorylated energy source \cite{BIT:BIT20026}. Finally, 10 mM phosphate was added to some reactions as described below in the form of potassium phosphate dibasic adjusted to pH 7.2 with acetic acid.

@@

The bounds on the transcription rate ($\mathcal{L}_{TX}=w_{TX}=\mathcal{U}_{TX}$) were modeled as:
\begin{equation}
	w_{TX} = V_{TX}^{max}\left(\frac{G}{K_{TX}+G}\right)
\end{equation}
where $G$ denotes the gene concentration and $K_{TX}$ denotes a transcription saturation coefficient.
The maximum rate of transcription $V_{TX}^{max}$ was formulated as:
\begin{equation}
	V_{TX}^{max} \equiv \left[R_{T}\left(\frac{v_{TX}}{l_{G}}\right)u\left(\kappa\right)\right]
\end{equation}
The term $R_{T}$ denotes the RNA polymerase concentration (nM),
$v_{TX}$ denotes the RNA polymerase elongation rate (nt/h),
$l_{G}$ denotes the gene length in nucleotides (nt).
The term $u\left(\kappa\right)$ (dimensionless, $0\leq u\left(\kappa\right)\leq 1$) describes a model of promoter activity, where $\kappa$ denotes promoter specific parameters.
The forms for the promoter models were taken from Moon $\emph{et~al.}$ \cite{Moon:2012ab}.
In this study, we considered two promoters: T7 and P70a.
The promoter function for the T7 promoter, $u_{T7}$, was given by:
\begin{equation}
	u_{T7} = \frac{K_{T7}}{1 + K_{T7}}
\end{equation}
where $K_{T7}$ denotes a T7 RNA polymerase binding constant.
The P70a promoter function $u_P70a$ (which was used for all other proteins) was formulated as:
\begin{equation}
	u_{P70a} = \frac{K_{1}+K_{2}f_{\sigma_{70}}}{1 + K_{1}+K_{2}f_{\sigma_{70}}}
\end{equation}
where $K_{1}$ denotes the weight of RNA polymerase binding alone,
$K_{2}$ denotes the weight of RNAP-$\sigma_{70}$ bound to the promoter,
and $f_{p70}$ denotes the fraction of the $\sigma_{70}$ transcription factor bound to RNAP, modeled as a Hill function: 
\begin{equation}
	f_{\sigma_{70}} = \frac{\sigma_{70}^{n}}{K_{D}^{n} + \sigma_{70}^{n}}
\end{equation}
where $\sigma_{70}$ denotes the sigma-factor 70 ($\sigma_{70}$) concentration, $K_{D}$ denotes the dissociation constant, and $n$ denotes the hill coefficient.
The values for all promoter parameters are given in Table ~\ref{tbl:parameters}.

The translation rate ($w_{TL}$) was bounded by:
 \begin{equation}
	0\leq w_{TL} \leq V_{TL}^{max}\left(\frac{mRNA_{SS}}{K_{TL}+mRNA_{SS}}\right)
\end{equation}
where $\rm mRNA_{SS}$ denotes the steady state mRNA abundance and $K_{TL}$ denotes the translation saturation constant.
The maximum translation rate $V_{TL}^{max}$ was formulated as:
\begin{equation}
	V_{TL}^{max} \equiv \left[K_{P} R_{X}\left(\frac{v_{TL}}{l_{P}}\right)\right]
\end{equation}
The term $K_{P}$ denotes the polysome amplification constant,
$v_{TL}$ denotes the ribosome elongation rate (amino acids per hour),
$l_{P}$ denotes the number of amino acids in the protein of interest,
and $\rm mRNA_{SS}$ denotes the steady-state mRNA concentration:
\begin{equation}
	 mRNA_{SS}\simeq\frac{w_{TX}}{\lambda}
\end{equation}
where $\lambda$ denotes the rate constant controlling the mRNA degradation rate.
All translation parameters are given in Table \ref{tbl:parameters}.
